---
title: "bvsModel"
output: html_document
---

This R markdown file experiments the proposed longitudinal Bayesian varaible selection model with several simulated datasets. The goal is to understand how the model performs under different settings.

```{r}
# requirements
require(MASS)
require(reshape2)
require(lme4)
require(lattice)
```

First, one needs to generate a control group. To be consistent, we use the estimated parameters from the mouse weight data to mimic its behaviors.

```{r}
set.seed(3)

#### parameter settings ####
days<- c(365,395,456,517,578,639,700,760,821,882,943,1004,1065,1125,1186)
W<- matrix(0, nrow=length(days), ncol=2)
W[,1]<- 1
W[,2]<- days/365 - 1 # standardize
Z<- W
X<- matrix(W[,2], ncol=1)
idmatrix<- diag(length(days))

#### prior settings ####
d1 = 75.95
d2 = 871.47
d3 = matrix(c(45.50, -5.75), ncol=1)
d4_inv = matrix(c(0.04, -0.02, -0.02, 0.06), nrow=2, ncol=2)

#### initial values ####
sigma2<- 5.06
lambdaD<- 1/((9.065+14.261)/2)
alpha<- matrix(c(45.50, -5.75), ncol=1)

#### generate control grp ####
n_contrl<- 2266
y_contrl<- matrix(0, nrow=n_contrl, ncol=length(days))
for(i in 1:n_contrl){
  y_contrl[i,]<- W%*%alpha +
                 Z%*%matrix(mvrnorm(1, matrix(0, nrow=2, ncol=1), diag(2)/rgamma(1, d1, d2)), ncol=1) +
                 matrix(mvrnorm(1, matrix(0, nrow=length(days), ncol=1), sigma2*idmatrix), ncol=1)
}

colnames(y_contrl)<- days

contrl_grp<- as.data.frame(y_contrl)
contrl_grp$id<- as.factor(seq(1, n_contrl))
contrl_grp$diet<- as.factor(99)

## convert to long format
contrl_grp<- melt(contrl_grp, id.vars = c('id', 'diet'), variable.name  = 'days', value.name = 'weight')
contrl_grp<- contrl_grp[order(contrl_grp$id, contrl_grp$days), ]
contrl_grp$days<- as.numeric(as.character(contrl_grp$days))
contrl_grp = contrl_grp[, c("days", "id", 'diet', 'weight')]

## save to file
#write.table(contrl_grp, "../simu_control_grp.txt", quote = FALSE, row.names=FALSE)
```

Let's look at the average of weights over all generated samples across time.

```{r, echo=FALSE}
plot(days, apply(y_contrl, 2, mean), 'o', col=2, main="Time vs. Mean Weight", xlab="Time", ylab="Mean Weight")
legend("topright", c("Control grp"), lty=1, col=2)
```

# Experiment No. 1

Generate a treatment group that starts at the same initial value as the control group with a set of different slopes (-5 to 5).

```{r}
set.seed(3)

simu_grp<- contrl_grp

id_idx<- n_contrl
diet_idx<- 1
## add Beta params to modify Alpha
for(beta in seq(-5, 5)){
  ## simulate data
  n_trt<- 1000
  y_trt<- matrix(0, nrow=n_trt, ncol=length(days))
  for(i in 1:n_trt){
    y_trt[i,]<- W%*%alpha + X*beta +
      Z%*%matrix(mvrnorm(1, matrix(0, nrow=2, ncol=1), diag(2)/rgamma(1, d1, d2)), ncol=1) +
      matrix(mvrnorm(1, matrix(0, nrow=length(days), ncol=1), sigma2*idmatrix), ncol=1)
    }
  
  colnames(y_trt)<- days
  
  trt_grp<- as.data.frame(y_trt)
  trt_grp$id<- as.factor(seq(id_idx+1, id_idx+n_trt))
  trt_grp$diet<- as.factor(diet_idx)
  
  id_idx = id_idx + n_trt
  diet_idx = diet_idx + 1
  
  ## convert to long format
  trt_grp<- melt(trt_grp, id.vars = c('id', 'diet'), variable.name  = 'days', value.name = 'weight')
  trt_grp<- trt_grp[order(trt_grp$id, trt_grp$days), ]
  trt_grp$days<- as.numeric(as.character(trt_grp$days))
  trt_grp = trt_grp[, c("days", "id", 'diet', 'weight')]
  
  simu_grp<- rbind(simu_grp, trt_grp)
}

## save to file
write.table(simu_grp, "../simu_grp1.txt", quote = FALSE, row.names=FALSE)
```

# Experiment No. 2

Generate a treatment group that starts at the same initial value as the control group with a set of different slopes (0 to 5).

```{r}
set.seed(3)

simu_grp<- contrl_grp

id_idx<- n_contrl
diet_idx<- 1
## add Beta params to modify Alpha
for(beta in seq(0, 5, 0.5)){
  ## simulate data
  n_trt<- 1000
  y_trt<- matrix(0, nrow=n_trt, ncol=length(days))
  for(i in 1:n_trt){
    y_trt[i,]<- W%*%alpha + X*beta +
      Z%*%matrix(mvrnorm(1, matrix(0, nrow=2, ncol=1), diag(2)/rgamma(1, d1, d2)), ncol=1) +
      matrix(mvrnorm(1, matrix(0, nrow=length(days), ncol=1), sigma2*idmatrix), ncol=1)
    }
  
  colnames(y_trt)<- days
  
  trt_grp<- as.data.frame(y_trt)
  trt_grp$id<- as.factor(seq(id_idx+1, id_idx+n_trt))
  trt_grp$diet<- as.factor(diet_idx)
  
  id_idx = id_idx + n_trt
  diet_idx = diet_idx + 1
  
  ## convert to long format
  trt_grp<- melt(trt_grp, id.vars = c('id', 'diet'), variable.name  = 'days', value.name = 'weight')
  trt_grp<- trt_grp[order(trt_grp$id, trt_grp$days), ]
  trt_grp$days<- as.numeric(as.character(trt_grp$days))
  trt_grp = trt_grp[, c("days", "id", 'diet', 'weight')]
  
  simu_grp<- rbind(simu_grp, trt_grp)
}

## save to file
write.table(simu_grp, "../simu_grp2.txt", quote = FALSE, row.names=FALSE)
```